---
title: "p8105_homework 3"
output: github_document
date: "2022-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(tidyverse)
library(lubridate)
```

## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

## Problem 2

### Loading and Cleaning Data
```{r}
accel_df = read_csv("./data/accel_data.csv") %>% janitor::clean_names() %>% pivot_longer(-c(week, day_id, day), names_to = "minute", values_to = "activity_counts") %>% mutate(minute = as.integer(gsub('[activity_]', '', minute))) %>% mutate(weekday_v_weekend = ifelse(day == "Saturday" | day == "Sunday", "Weekend", "Weekday")) %>% mutate(day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))
```

In order to clean the data, we first cleaned the named using `janitor::clean_names()`. We then used `pivot_longer` to tidy the data and create a row for each observation (each minute was given an entry in the newly created `minute` variable). The corresponding activity count values were added to the `activity_counts` variable. We then removed the "activity_" prefix in the `minute` variable to have a integer variable that only contained the numeric values of the minutes. A `weekday_v_weekend` variable was created to distinguish between the types of days. The `day` variable was coerced into a factor variable based on the days of the week, starting with Sunday. The resulting dataset contains `r nrow(accel_df)` observations (one minute for each of the 35 days). The dataset contains the following variables: `week` to represent the week number, `day_id` to represent the day number (ranging from 1 to 35), `day` variable to represent the day of the week, `minute` to represent which minute of the day, `activity_counts` to represent the activity at that specific minute, and `weekday_v_weekend` to represent the type of day. The mean activity count per minute was `r format(mean(accel_df$activity_counts), scientific = F, digits = 6)`. 

### Total Activity Over the Days
We will now aggregate across minutes and create a table showing the total activity per day.
```{r, message = FALSE}
total_activity_df = accel_df %>%
  group_by(day_id, day) %>%
  summarize(total_activity = sum(activity_counts))

total_activity_df %>%
  knitr::kable(digits = 2)
```

It can be noted that, according to the table, the man's activity levels were the lowest on the fourth and fifth Saturdays. On these days, his activity count were valued at 1440. These values were much lower than the mean activity level, which was approximately `r format(mean(total_activity_df$total_activity), scientific = F, digits = 6)`. In order to study the patterns of activity level each day more effectively, we will create a scatterplot using `day_id` as our x-axis and `total_activity` as our y-axis. We will use `geom_smooth` to estimate a conditional mean to study how the the activity level changes across the 5 weeks.

```{r, message = FALSE}
ggplot(total_activity_df, aes(x = day_id, y = total_activity)) + 
  geom_point(aes(color = day)) +
  geom_smooth(se = FALSE) +  labs(
    title = "Total Activity Plot",
    x = "Day ID",
    y = "Total Activity Per Day",
  )
```
Viewing the plot, it is apparent that the individual was gradually increasing his activity levels until the second week, after which his activity levels decreased. This continued until around the fourth week when his activity levels began to increase again. The plot also confirms that the individual had his lowerst activity levels on the fourth and fifth Saturdays.

### Activity Over the Course of the Day
```{r, message = FALSE}
accel_hour_df = accel_df %>% 
  mutate(hour = floor(minute/60)) %>% 
  group_by(hour, day, day_id) %>% 
  summarize(hourly_total = sum(activity_counts)) %>% arrange(day_id)

ggplot(accel_hour_df, aes(x = hour, y = hourly_total)) + 
  geom_point(aes(color = day)) +
  geom_smooth(se = FALSE) +  labs(
    title = "Activity Over the Course of the Day",
    x = "Hour (military time)",
    y = "Activity")
```

Overall, it appears that the individual 

## Problem 2

```{r}
library(p8105.datasets)
data("ny_noaa")
```

### Write Description

```{r}
ny_noaa_df = ny_noaa %>% separate(col = date, into = c("year", "month", "day")) %>% mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>% mutate(prcp = prcp/100, snow = snow/100, snwd = snwd/100, tmax = as.numeric(tmax)/10, tmin = as.numeric(tmin)/10)

```
The most frequent value was `r `. This is due to the fact that many stations did not report a value for snowfall. Removing the missing snowfall values, we find that the mode is `r `